---
title: "Practical Machine Learning Course Project"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading of needed libraries
First load required libraries/variables/parameters needed for this project:
```{r}
library(ggplot2)
library(lattice)
library(caret)
library(plyr)
library(parallel)
library(foreach)
library(iterators)
library(doParallel)
library(randomForest)
knitr::opts_chunk$set(echo=TRUE, fig.path='Figures/', fig.width=10, fig.height=4, cache=TRUE)
```

## Preparing/getting the data
First, check if the file "activity.zip" exists in the repository. If not, download it. Then unzip and load data into data frame `data`. 
```{r}
if(!file.exists("pml-training.csv")) {
        download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv","pml-training.csv")
}
train_data <- read.csv("pml-training.csv")

if(!file.exists("pml-testing.csv")) {
        download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv","pml-testing.csv")
}
test_data <- read.csv("pml-testing.csv")
tr_dim <- dim(train_data)
te_dim <- dim(test_data)
```
The training data set has `r tr_dim[1]` observations and `r tr_dim[2]` variables. The testing data, on the other hand, has `r te_dim[1]` observations and `r te_dim[2]` variables.

## Cleaning the data 
Some cleaning has to be done to the data before a model can be built. Missing values should be removed. Also, columns that are not useful/contribute much to the accelerometer values can be removed.

```{r}
train_data <- train_data[, colSums(is.na(train_data)) == 0]
test_data <- test_data[, colSums(is.na(test_data)) == 0]
classe <- train_data$classe
trainTemp <- grepl("^X|timestamp|window", names(train_data))
train_data <- train_data[, !trainTemp]
train_data <- train_data[, sapply(train_data, is.numeric)]
train_data$classe <- classe
testTemp <- grepl("^X|timestamp|window", names(test_data))
test_data <- test_data[, !testTemp]
test_data <- test_data[, sapply(test_data, is.numeric)]
tr_dim <- dim(train_data)
te_dim <- dim(test_data)
```
After cleaning, the training data set has `r tr_dim[1]` observations and `r tr_dim[2]` variables. The testing data now has `r te_dim[1]` observations and `r te_dim[2]` variables. "classe" is the outcome/label/class variable.

## Modeling the data
The cleaned training data can be partitioned into 2 parts - 70% of the data is the training data set, and the rest 30% will be used as the validation data set. Random Forest will be selected for the predictive model. We will use a 10-fold cross-validation technique to evaluate the algorithm.

```{r}
set.seed(23232)
ptm <- proc.time()
inTrain <- createDataPartition(train_data$classe, p=0.70, list=F)
training <- train_data[inTrain,]
testing <- train_data[-inTrain,]

controlRF <- trainControl(method="cv", 10)
modelRF <- train(classe ~ ., data=training, method="rf", trControl=controlRF, ntree = 250)
modelRF
predictRF <- predict(modelRF, testing)
confusionMatrix(testing$classe, predictRF)

accuracy <- postResample(predictRF, testing$classe)
accuracy
oose <- 1- as.numeric(confusionMatrix(testing$classe, predictRF)$overall[1])
oose
```